---
title: "RPDRage"
author: "Danni Brower"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Age of Drag Race: Is Rupaul's Drag Race biased towards younger queens? 

&nbsp;

### Research Question

Rupaul's Drag Race is a reality competition show for drag queens in which contestants compete to be America's next Drag Super Star. Amongst the many topics of debate surrounding the show is whether it has a bias towards young queens.  



This project visualises the spread of the ages of Drag Race contestants by season, as well as the ages of the queens that won each season. This will allow a general picture of the age range of the queens as well as the identification of whether there is a historical trend in it.

&nbsp;

### Data Origins

This data is provided as part of the package *dragracer*^1^. The CRAN archive (?) for the package is <https://cran.r-project.org/package=dragracer>.

A description of the project is available at <http://svmiller.com/blog/2019/02/dragracer-rupauls-drag-race-analysis/>.
The data was compiled by scraping the Rupaul's Drag Race Wiki: <https://rupaulsdragrace.fandom.com/wiki/RuPaul%27s_Drag_Race_Wiki>

Because it is a package, the first time you run this script, you must run ``install.packages(dragracer)`` first.

The *dragracer* package includes three datasets:
*rpdr_contestants*, which provides information about the contestants,
*rpdr_contep*, which describes each contestant's performance in each episode,
*rpdr_ep*, which gives characteristics of each episode. This analysis and visualisation use the *rdpr_contestants* and *rpdr_contep* datasets.

The relevant datasets have been exported to csv files using this code: ``write.csv(df, "filepath\\filename.csv", row.names = FALSE)``. They are available in the data folder.

```{r dragracer packages, results = 'hide', message = FALSE, warning=FALSE}
#loading required packages, including the datasets

library(ggplot2)
library(plotly)
library(magrittr)
library(tidyverse)
```


```{r dragracer datasets, warning=FALSE}
library(dragracer) #when loaded returns a catchphrase from the show
```


```{r dragracer preview}
#previewing datasets

head(rpdr_contestants)
head(rpdr_contep)
```

## Data Wrangling

```{r data wrangling}

#create a data frame including the contestant's ages, season, and final ranking in the competition

combined <- left_join(rpdr_contestants, rpdr_contep)

ranked <- combined %>% 
          select(c("season", "contestant","rank", "age")) %>%
          distinct() %>%
          group_by("season")

#create a data frame of winners only

winners <- subset(ranked, rank == "1")

head(ranked)

#simple statistics

mean(ranked$age)

mean(winners$age)




```

## Visualisation

```{r age plot, echo=TRUE}

# violin and scatterplots of all contestants' ages

age_all <- ggplot(ranked, aes(x = factor(season), y = age)) +
    geom_violin(aes(color = season, fill = season), alpha = 0.5) +
    stat_summary(fun = "mean",
                 geom = "crossbar", 
                 width = 0.3) +
    geom_point(aes(color = season, fill = season),
                   position = position_jitterdodge(jitter.width = .1, dodge.width = 0), 
                   shape=21, size = 1) +
  
# graph aesthetics   
  
    labs(x = element_blank()) +
    ggtitle("Rupaul's Drag Race Contestant Ages") +
    theme(legend.position="none",  
    strip.background=element_blank(),
    axis.text.x=element_text(size=10,color="black"),
    axis.text.y=element_text(size=10,color="black"),
    panel.background = element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.border = element_rect(size= .75,fill = NA,color = "black"),
    plot.margin = unit(c(.5, 0.5, .5, .5), "cm"))
  
#adding interactive overlay of winners' ages
#clean up code?

age_plot <- ggplotly(age_all + geom_point(data = winners, aes(x = factor(season))))

age_plot <- age_plot %>% style(hoverinfo = "skip", traces = 1:40, text = winners$age)
age_plot <- age_plot %>% style(text = paste("Winner's age:", winners$age))

age_plot

```

## Summary, Discussion

Drag Race contestants have ranged in age from 21 to 52. The average age of all participants through season 13 is `r round(mean(ranked$age), digits = 0)`. Winners average age is `r round(mean(winners$age), digits = 0)` There does not seem to be a significant trend in age across seasons. The relatively young average age of all contestants and winners upholds the commonly accepted belief that drag is a young girl's game. As beloved Drag Race contestant Katya Zamolodchkava has said, "There's nothing more depressing than an older drag queen that doesn't want to do it."^(2)^


&nbsp;
&nbsp;
&nbsp;
&nbsp;


**References**

1. Steve Miller (2021). dragracer: Data Sets for RuPaul's Drag Race. R package version 0.1.6. https://CRAN.R-project.org/package=dragracer
2. Mattel, Trixie & Zamolodchikova, Katya. (2020, Feb 12). UNHhhh Ep. 114: I'm in My 30s. Youtube. https://www.youtube.com/watch?v=OilvWfWkaaw


&nbsp;
&nbsp;
&nbsp;
&nbsp;

